### 2017-09-21 sheets/talk-6/sheet-6.tex

\begin{answer}
  Let $(A, \le)$ be a poset concidered as a category and $I$ a set.
  A \demph{lower bound} of a family $(x_i)_{i\in I}$ is an element $a \in A$, s.\,t. $\forall i \in I: a \le x_i$.
  $z \in A$ is a \demph{greatest lower bound} of this family if $\forall a \in A: \text{$a$ is a lower bound} \Rightarrow a \le z$.

@@

\begin{definition}{Minus-core of linear maps in vector spaces}
  Let $V, W$ be vector spaces over a field $K$. Let $s, t \from V \to W$ be linear functions.
  We define the linear function $s - t := \lambda v\in V. \; s(v) - t(v)$.
  Remember that for $g \from V \to W$ we have defined $core(g) := \set{ v \in V | g(v) = 0 }$.
  Now $core (s - t) = \set{v \in V | s(v) = t (v) }$.
\end{definition}

\begin{exercise}[Equalizer of vector spaces]
  Let
  $\xymatrix{
      X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
  }$
  be a diagram in $Vect_K$ for a field $K$. Show that $core(s-t)$ is the equalizer for this diagram.
\end{exercise}

@@

\begin{answer}
  \begin{proof}
    We show that $S := P$ is the coproduct of $X$ and $Y$ together with the coprojection functions $p_1$ and $p_2$.
    Let
    $ \xymatrix{
      X \ar[r]^{f_1} & A & Y \ar[l]_{f_2}
    } $
    be a diagram in $\cat{A}$.
    We have to show that there is a unique map $\bar{f}' \from P \to A$ such that
    \[ \xymatrix{
      X \ar[rdd]_{f_1}\ar[rd]^{p_1} & & Y \ar[ldd]^{f_2} \ar[ld]_{p_2} \\
      & P \ar@{.>}[d]|{\bar{f}'} & \\
      & A &
    } \]
    commutes. We know that the diagram
    \[ \xymatrix{
      Z \ar[r]_s \ar[d]^t & X \ar[d]_{f_1} \\
      Y \ar[r]^{f_2} & A
    } \]
    commutes because $Z$ is initial.
    Now because this square is then a pushout square there exists a unique map $\bar{f}\from P \to A$ such that
    \[ \xymatrix{
      Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \ar[ddr]^{f_1} & \\
      Y \ar[r]^{p_2} \ar[drr]_{f_2}& P \ar@{.>}[dr]|{\bar{f}} & \\
      & & A  \\
    } \]
    commutes. We see that we can choose $\bar{f}' := \bar{f}$.
    Suppose we have another $\tilde{f} \from P \to A$ such that
    \[ \xymatrix{
      X \ar[rdd]_{f_1}\ar[rd]^{p_1} & & Y \ar[ldd]^{f_2} \ar[ld]_{p_2} \\
      & P \ar@{.>}[d]|{\tilde{f}} & \\
      & A &
    } \]
    commutes. Then
    \[ \xymatrix{
      Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \ar[ddr]^{f_1} & \\
      Y \ar[r]^{p_2} \ar[drr]_{f_2}& P \ar@{.>}[dr]|{\tilde{f}} & \\
      & & A  \\
    } \]
    would also commute (because $Z$ is initial) but $\bar{f}$ is unique, hence $\bar{f} = \tilde{f}$.
  \end{proof}

@@

\begin{answer}
  \begin{proof}
    We give names to the arrows and objects:
    \[ \xymatrix{
      A \ar[r]^h \ar[d]^j & B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
      D \ar[r]^m & E \ar[r]^n & F
    } \]
    Now let the pullback square
    \[ \xymatrix{
      B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
      E \ar[r]^n & F
    } \]
    be the pullback of $\xymatrix{E \ar[r]^n & F & C \ar[l]_l}$.
    We show both directions of the equivalence:
    \begin{itemize}
      \item[``$\Rightarrow$'']
        Let the pullback square
        \[ \xymatrix{
          A \ar[r]^{i \of h} \ar[d]^j & C \ar[d]^l \\
          D \ar[r]^{n \of m} & F
        } \]
        be the pullback of $\xymatrix{D \ar[r]^{n \of m} & F & C \ar[l]_l}$.
        Let
        \[ \xymatrix{
          Z \ar[r]^{f_1} \ar[d]^{f_2} & B \ar[d]^k \\
          D \ar[r]^{m} & E
        } \]
        be another pullback square for the left-hand square. Now
        \[ \xymatrix{
          Z \ar[r]^{f_1} \ar[d]^{f_2} & B \ar[r]^{i} \ar[d]^k & C \ar[d]^l \\
          D \ar[r]^{m} & E \ar[r]^n & F
        } \]
        commutes. Now because the outer rectangle is a pullback there is a unique map $\bar{f} \from Z \to A$, such that
        \[ \xymatrix{
          Z \ar@/^/[drrr]^{i \of f_1}
            \ar@{.>}[dr]|{\bar{f}}
            \ar@/_/[ddr]_{f_2} & & \\
            & A \ar[r]_h \ar[d]^j & B \ar[r]_{i} & C \ar[d]_l \\
            & D \ar[r]^{m} & E \ar[r]^n & F
        } \]
        commutes. We show that this $\bar{f}$ is the unique map such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{f_1}
            \ar@{.>}[dr]|{\bar{f}}
            \ar@/_/[ddr]_{f_2} \\
          & A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
          & D \ar[r]^m & E
        } \]
        commutes. We have to show $j \of \bar{f} = f_2$ and $h \of \bar{f} = f_1$. The first equatation holds by assumption.

        Now we define two maps
        \begin{alignat*}{4}
          g_1 &\from Z \to C&  \qquad g_1 &:= i \of f_1 \\
          g_2 &\from Z \to E&         g_2 &:= m \of j \of \bar{f}
        \end{alignat*}
        and show that the diagram
        \[ \xymatrix{
          Z \ar[r]^{g_1} \ar[d]^{g_2} & C \ar[d]^l \\
          E \ar[r]^n & F
        } \]
        commutes and is therefore a pullback square of the right-hand square:
        \begin{alignat*}{2}
          l \of g_1
          &= l \of i \of f_1 \\
          &= l \of i \of h \of \bar{f} \\
          &= n \of m \of j \of \bar{f} \\
          &= n \of g_2
        \end{alignat*}
        Now because the right-hand square is a pullback there is a unique map $\bar{g} \from Z \to B$, such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{g_1}
            \ar@{.>}[dr]|{\bar{g}}
            \ar@/_/[ddr]_{g_2} \\
          & B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
          & E \ar[r]^n & F
        } \]
        commutes.
        By showing that this diagram comutes for $\bar{g} = f_1$ and for $\bar{g} = h \of \bar{f}$ we show $h \of \bar{f} = f_2$.
        \begin{itemize}
          \item[1.]$\bar{g} = f_1$: We have $i \of f_1 = g_1$ by definition and $k \of f_1 = m \of f_2 = m \of j \of \bar{f} = g_2$.
          \item[2.]$\bar{g} = h \of \bar{f}$: We have $(i \of h) \of \bar{f} = i \of f_2$ and $(k \of h) \of \bar{f} = (m \of j) \of \bar{f}$.
        \end{itemize}

@@

      \item[``$\Leftarrow$'']
        Let the pullback square
        \[ \xymatrix{
          A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
          D \ar[r]^m & E
        } \]
        be the pullback of $\xymatrix{D \ar[r]^m & E & B \ar[l]_k}$.
        Let
        $ \xymatrix{
          D & Z \ar[l]_{f_2} \ar[r]^{f_1} & C
        } $
        be a diagram such that
        \[ \xymatrix{
          Z \ar[r]^{f_1} \ar[d]_{f_2} & C \ar[d]^l \\
          D \ar[r]^{n\of m} & F
        } \]
        commutes. We note that
        \[ \xymatrix{
          Z \ar[r]^{f_1} \ar[d]_{m \of f_2} & C \ar[d]^l \\
          E \ar[r]_n & F
        } \]
        is thus a pullback square of the right-hand square.
        Thus we have a unique map $\bar{f} \from Z \to B$ such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{f_1}
            \ar@{.>}[dr]|{\bar{f}}
            \ar@/_/[ddr]_{m \of f_2} \\
          & B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
          & E \ar[r]^n & F
        } \]
        commutes. Then
        \[ \xymatrix{
          Z \ar[r]^{\bar{f}} \ar[d]^{f_2} & B \ar[d]^k \\
          D \ar[r]^m && E
        } \]
        is a pullback square of the left-hand square.
        Thus we have a unique map $\bar{g} \from Z \to A$ such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{\bar{f}}
            \ar@{.>}[dr]|{\bar{g}}
            \ar@/_/[ddr]_{f_2} \\
          & A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
          & D \ar[r]^m & E
        } \]
        commutes. We show that this $\bar{g}$ is the unique map such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{f_1}
            \ar@{.>}[dr]|{\bar{g}}
            \ar@/_/[ddr]_{f_2} \\
          & A \ar[r]^{i \of h} \ar[d]^j & C \ar[d]^l \\
          & D \ar[r]^{n \of m} & F
        } \]
        commutes.
        The diagram comutes for $\bar{g}$:
        \begin{align*}
          i \of h \of \bar{g} = i \of \bar{f} = f_1 \\
          j \of \bar{g} = f_2
        \end{align*}
        Let $\tilde{g}$ be another map such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{f_1}
            \ar@{.>}[dr]|{\tilde{g}}
            \ar@/_/[ddr]_{f_2} \\
          & A \ar[r]^{i \of h} \ar[d]^j & C \ar[d]^l \\
          & D \ar[r]^{n \of m} & F
        } \]
        commutes. To show $\bar{g} = \tilde{g}$ it is because of the uniqueness enough to show that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{\bar{f}}
            \ar@{.>}[dr]|{\tilde{g}}
            \ar@/_/[ddr]_{f_2} \\
          & A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
          & D \ar[r]^m & E
        } \]
        commutes.
        We already have $j \of \tilde{g} = f_2$ and only have to show $h \of \tilde{g} = \bar{f}$.
        Because of $\bar{f} = h \of \bar{g}$ it suffices to show $h \of \tilde{g} = h \of \bar{g}$.
        The commuting square
        \[ \xymatrix{
          Z \ar[r]^{i \of h \of \tilde{g}} \ar[d]_{m \of j \of \tilde{g}} & C \ar[d]^l \\
          E \ar[r]^n & F
        } \]
        is another pullback square for the right-hand square.
        Thus we have a unique map $\bar{x} \from Z \to B$ such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{i \of h \of \tilde{g}}
            \ar@{.>}[dr]|{\bar{x}}
            \ar@/_/[ddr]_{m \of j \of \tilde{g}} \\
          & B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
          & E \ar[r]^n & F
        } \]
        commutes.
        We get $h \of \tilde{g} = h \of \bar{g}$ when we show that this diagram commutes for $\bar{x} = h \of \tilde{g}$ and $\bar{x} = h \of \bar{g}$.

>>>>

\begin{answer}
  Let $(A, \le)$ be a poset considered as a category and $I$ a set.
  A \demph{lower bound} of a family $(x_i)_{i\in I}$ is an element $a \in A$, s.\,t. $\forall i \in I: a \le x_i$.
  $z \in A$ is a \demph{greatest lower bound} of this family if $\forall a \in A: \text{$a$ is a lower bound} \Rightarrow a \le z$.

@@

\begin{definition}{Minus-core of linear maps between vector spaces}
  Let $V, W$ be vector spaces over a field $K$. Let $s, t \from V \to W$ be linear functions.
  We define the linear function $s - t := \lambda v\in V. \; s(v) - t(v)$.
  Remember that for $g \from V \to W$ we have defined $core(g) := \set{ v \in V | g(v) = 0 }$.
  Now $core (s - t) = \set{v \in V | s(v) = t (v) }$.
\end{definition}

\begin{exercise}[Equalizer of vector spaces]
  Let
  $\xymatrix{
      X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
  }$
  be a diagram in $\Vect_K$ for a field $K$. Show that $core(s-t)$ is the equalizer for this diagram.
\end{exercise}

@@

\begin{answer}
  \begin{proof}
    We show that $P$ is the coproduct of $X$ and $Y$ together with the coprojection functions $p_1$ and $p_2$.
    Let
    $ \xymatrix{
      X \ar[r]^{f_1} & A & Y \ar[l]_{f_2}
    } $
    be a diagram in $\cat{A}$.
    We have to show that there is a unique map $\bar{f}' \from P \to A$ such that
    \[ \xymatrix{
      X \ar[rdd]_{f_1}\ar[rd]^{p_1} & & Y \ar[ldd]^{f_2} \ar[ld]_{p_2} \\
      & P \ar@{.>}[d]|{\bar{f}'} & \\
      & A &
    } \]
    commutes. We know that the diagram
    \[ \xymatrix{
      Z \ar[r]_s \ar[d]^t & X \ar[d]_{f_1} \\
      Y \ar[r]^{f_2} & A
    } \]
    commutes because $Z$ is initial.
    Now because this square is then a pushout square there exists a unique map $\bar{f}\from P \to A$ such that
    \[ \xymatrix{
      Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \ar[ddr]^{f_1} & \\
      Y \ar[r]^{p_2} \ar[drr]_{f_2}& P \ar@{.>}[dr]|{\bar{f}} & \\
      & & A  \\
    } \]
    commutes. We see that we can choose $\bar{f}' := \bar{f}$.
    Suppose we have another $\tilde{f} \from P \to A$ such that
    \[ \xymatrix{
      X \ar[rdd]_{f_1}\ar[rd]^{p_1} & & Y \ar[ldd]^{f_2} \ar[ld]_{p_2} \\
      & P \ar@{.>}[d]|{\tilde{f}} & \\
      & A &
    } \]
    commutes. Then
    \[ \xymatrix{
      Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \ar[ddr]^{f_1} & \\
      Y \ar[r]^{p_2} \ar[drr]_{f_2}& P \ar@{.>}[dr]|{\tilde{f}} & \\
      & & A  \\
    } \]
      would also commute (because $Z$ is initial), but $\bar{f}$ is unique, hence $\bar{f} = \tilde{f}$.
  \end{proof}

@@

\begin{answer}
  \begin{proof}
    We give names to the arrows and objects:
    \[ \xymatrix{
      A \ar[r]^h \ar[d]^j & B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
      D \ar[r]^m & E \ar[r]^n & F
    } \]
    Now let the pullback square
    \[ \xymatrix{
      B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
      E \ar[r]^n & F
    } \]
    be the pullback of $\xymatrix{E \ar[r]^n & F & C \ar[l]_l}$.
    We show both directions of the equivalence:
    \begin{itemize}
      \item[``$\Rightarrow$'']
        Let the pullback square
        \[ \xymatrix{
          A \ar[r]^{i \of h} \ar[d]^j & C \ar[d]^l \\
          D \ar[r]^{n \of m} & F
        } \]
        be the pullback of $\xymatrix{D \ar[r]^{n \of m} & F & C \ar[l]_l}$.
        Let
        \[ \xymatrix{
          Z \ar[r]^{f_1} \ar[d]^{f_2} & B \ar[d]^k \\
          D \ar[r]^{m} & E
        } \]
        be another pullback square for the left-hand square. Now
        \[ \xymatrix{
          Z \ar[r]^{f_1} \ar[d]^{f_2} & B \ar[r]^{i} \ar[d]^k & C \ar[d]^l \\
          D \ar[r]^{m} & E \ar[r]^n & F
        } \]
        commutes. Now because the outer rectangle is a pullback there is a unique map $\bar{f} \from Z \to A$, such that
        \[ \xymatrix{
          Z \ar@/^/[drrr]^{i \of f_1}
            \ar@{.>}[dr]|{\bar{f}}
            \ar@/_/[ddr]_{f_2} & & \\
            & A \ar[r]_h \ar[d]^j & B \ar[r]_{i} & C \ar[d]_l \\
            & D \ar[r]^{m} & E \ar[r]^n & F
        } \]
        commutes. We show that this $\bar{f}$ is the unique map such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{f_1}
            \ar@{.>}[dr]|{\bar{f}}
            \ar@/_/[ddr]_{f_2} \\
          & A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
          & D \ar[r]^m & E
        } \]
        commutes. We have to show $j \of \bar{f} = f_2$ and $h \of \bar{f} = f_1$. The first equation holds by assumption.

        Now we define two maps
        \begin{alignat*}{4}
          g_1 &\from Z \to C&  \qquad g_1 &:= i \of f_1 \\
          g_2 &\from Z \to E&         g_2 &:= m \of j \of \bar{f}
        \end{alignat*}
        and show that the diagram
        \[ \xymatrix{
          Z \ar[r]^{g_1} \ar[d]^{g_2} & C \ar[d]^l \\
          E \ar[r]^n & F
        } \]
        commutes and is therefore a pullback square of the right-hand square:
        \begin{alignat*}{2}
          l \of g_1
          &= l \of i \of f_1 \\
          &= l \of i \of h \of \bar{f} \\
          &= n \of m \of j \of \bar{f} \\
          &= n \of g_2
        \end{alignat*}
        Now because the right-hand square is a pullback there is a unique map $\bar{g} \from Z \to B$, such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{g_1}
            \ar@{.>}[dr]|{\bar{g}}
            \ar@/_/[ddr]_{g_2} \\
          & B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
          & E \ar[r]^n & F
        } \]
        commutes.
        By showing that this diagram commutes for $\bar{g} = f_1$ and for $\bar{g} = h \of \bar{f}$ we show $h \of \bar{f} = f_2$.
        \begin{itemize}
          \item[1.]$\bar{g} = f_1$: We have $i \of f_1 = g_1$ by definition and $k \of f_1 = m \of f_2 = m \of j \of \bar{f} = g_2$.
          \item[2.]$\bar{g} = h \of \bar{f}$: We have $(i \of h) \of \bar{f} = i \of f_2$ and $(k \of h) \of \bar{f} = (m \of j) \of \bar{f}$.
        \end{itemize}

@@

      \item[``$\Leftarrow$'']
        Let the pullback square
        \[ \xymatrix{
          A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
          D \ar[r]^m & E
        } \]
        be the pullback of $\xymatrix{D \ar[r]^m & E & B \ar[l]_k}$.
        Let
        $ \xymatrix{
          D & Z \ar[l]_{f_2} \ar[r]^{f_1} & C
        } $
        be a diagram such that
        \[ \xymatrix{
          Z \ar[r]^{f_1} \ar[d]_{f_2} & C \ar[d]^l \\
          D \ar[r]^{n\of m} & F
        } \]
        commutes. We note that
        \[ \xymatrix{
          Z \ar[r]^{f_1} \ar[d]_{m \of f_2} & C \ar[d]^l \\
          E \ar[r]_n & F
        } \]
        is thus a pullback square of the right-hand square.
        Thus we have a unique map $\bar{f} \from Z \to B$ such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{f_1}
            \ar@{.>}[dr]|{\bar{f}}
            \ar@/_/[ddr]_{m \of f_2} \\
          & B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
          & E \ar[r]^n & F
        } \]
        commutes. Then
        \[ \xymatrix{
          Z \ar[r]^{\bar{f}} \ar[d]^{f_2} & B \ar[d]^k \\
          D \ar[r]^m && E
        } \]
        is a pullback square of the left-hand square.
        Thus we have a unique map $\bar{g} \from Z \to A$ such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{\bar{f}}
            \ar@{.>}[dr]|{\bar{g}}
            \ar@/_/[ddr]_{f_2} \\
          & A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
          & D \ar[r]^m & E
        } \]
        commutes. We show that this $\bar{g}$ is the unique map such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{f_1}
            \ar@{.>}[dr]|{\bar{g}}
            \ar@/_/[ddr]_{f_2} \\
          & A \ar[r]^{i \of h} \ar[d]^j & C \ar[d]^l \\
          & D \ar[r]^{n \of m} & F
        } \]
        commutes.
        The diagram commutes for $\bar{g}$:
        \begin{align*}
          i \of h \of \bar{g} = i \of \bar{f} = f_1 \\
          j \of \bar{g} = f_2
        \end{align*}
        Let $\tilde{g}$ be another map such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{f_1}
            \ar@{.>}[dr]|{\tilde{g}}
            \ar@/_/[ddr]_{f_2} \\
          & A \ar[r]^{i \of h} \ar[d]^j & C \ar[d]^l \\
          & D \ar[r]^{n \of m} & F
        } \]
        commutes. To show $\bar{g} = \tilde{g}$ it is because of the uniqueness enough to show that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{\bar{f}}
            \ar@{.>}[dr]|{\tilde{g}}
            \ar@/_/[ddr]_{f_2} \\
          & A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
          & D \ar[r]^m & E
        } \]
        commutes.
        We already have $j \of \tilde{g} = f_2$ and only have to show $h \of \tilde{g} = \bar{f}$.
        Because of $\bar{f} = h \of \bar{g}$ it suffices to show $h \of \tilde{g} = h \of \bar{g}$.
        The commuting square
        \[ \xymatrix{
          Z \ar[r]^{i \of h \of \tilde{g}} \ar[d]_{m \of j \of \tilde{g}} & C \ar[d]^l \\
          E \ar[r]^n & F
        } \]
        is another pullback square for the right-hand square.
        Thus we have a unique map $\bar{x} \from Z \to B$ such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{i \of h \of \tilde{g}}
            \ar@{.>}[dr]|{\bar{x}}
            \ar@/_/[ddr]_{m \of j \of \tilde{g}} \\
          & B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
          & E \ar[r]^n & F
        } \]
        commutes.
        We get $h \of \tilde{g} = h \of \bar{g}$ when we show that this diagram commutes for $\bar{x} = h \of \tilde{g}$ and $\bar{x} = h \of \bar{g}$.
